Draft

### 2022-06-15 draft/draft.qmd
finished eda and method writeup

Before we select our predictor variables, we create a visualization and
summary statistics for the response variable.

@@

We now conduct exploratory data analysis (EDA) on each of the 15
predictor variables in the data set. The EDA for each variable comprises
a histogram and a simple linear regression (SLR) against the outcome.

@@

```{r eda-age}
#| warning: false
#Age
abortion_data <- abortion_data %>% 
  mutate(agegroup = case_when(age >= 90 ~ '9',
                                             age >= 80  & age <= 89 ~ '8',
                                             age >= 70  & age <= 79 ~ '7',
                                             age >= 60  & age <= 69 ~ '6',
                                             age >= 50  & age <= 59 ~ '5',
                                             age >= 40  & age <= 49 ~ '4',
                                             age >= 30  & age <= 39 ~ '3',
                                             age >= 20  & age <= 29 ~ '2',
                              age <= 19 ~ '1'))
abortion_data
ggplot(abortion_data, aes(x = age)) +
  geom_histogram(binwidth = 1) + 
  labs(title = "Histogram of age")
ggplot(abortion_data, aes(x = agegroup, group = agegroup, y = aj)) +
  geom_boxplot()
```

```{r eda-college}
#| warning: false
# College education
ggplot(abortion_data, aes(x = collegeed)) +
  geom_histogram() + 
  labs(title = "Histogram of College education")
ggplot(abortion_data, aes(x = collegeed, group = collegeed, y = aj)) +
  geom_boxplot()
```

```{r eda-female}
#| warning: false
# Female
ggplot(abortion_data, aes(x = female)) +
  geom_histogram() + 
  labs(title = "Histogram of female")
ggplot(abortion_data, aes(x = female,group = female, y = aj)) +
  geom_boxplot()
```

```{r eda-unemployed}
#| warning: false
# Unemployed
ggplot(abortion_data, aes(x = unemployed)) +
  geom_histogram() + 
  labs(title = "Histogram of unemployed")
ggplot(abortion_data, aes(x = unemployed, group = unemployed, y = aj)) +
  geom_boxplot()
```

@@

```{r eda-financial}
#| warning: false
# Financial satisfaction
ggplot(abortion_data, aes(x = satisfinancial)) +
  geom_histogram() + 
  labs(title = "Histogram of Financial satisfaction")
ggplot(abortion_data, aes(x = satisfinancial,group = satisfinancial, y = aj)) +
  geom_boxplot()
```

```{r eda-materialism}
#| warning: false
# WVS post-materialist index
ggplot(abortion_data, aes(x = postma4)) +
  geom_histogram() + 
  labs(title = "Histogram of WVS post-materialist index")
ggplot(abortion_data, aes(x = postma4, group = postma4,y = aj)) +
  geom_boxplot()
```

@@

```{r eda-trust}
#| warning: false
# Trust most people
ggplot(abortion_data, aes(x = trustmostpeople)) +
  geom_histogram() + 
  labs(title = "Histogram of whether respondent trusted most people")
ggplot(abortion_data, aes(x = trustmostpeople, group = trustmostpeople, y = aj)) +
  geom_boxplot()
```

@@

```{r pair1}
#| warning: false
#I am doing ggpairs on 13 predictor variables. 
#To ensure they fit into the screen, I will make 4 matrices with subsets

ggpairs(abortion_data, 
  columns = c("year", "age", "collegeed", "female", "unemployed",
              "ideology", "satisfinancial"),
  columnLabels = c("year", "age", "college", "female", 
                   "unemployed", "ideology", "fin satis.")) +
   theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(angle = 45, size = 10),
    strip.text.y = element_text(angle = 0, hjust = 0)
    )
```

```{r pair2}
#| warning: false
ggpairs(abortion_data, 
  columns = c("postma4", "cai", "trustmostpeople", "godimportant", 
              "respectauthority","nationalpride"),
  columnLabels = c("postmaterialism", "child autonomy", "trust", "god", 
              "respectauthority","nationalpride")) +
   theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(angle = 45, size = 10),
    strip.text.y = element_text(angle = 0, hjust = 0)
    )
```

```{r pair3}
#| warning: false
ggpairs(abortion_data, 
  columns = c("year", "age", "collegeed", "female", 
              "respectauthority","nationalpride"),
  columnLabels = c("year", "age", "collegeed", "female", 
              "respectauthority","nationalpride")) +
   theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(angle = 45, size = 10),
    strip.text.y = element_text(angle = 0, hjust = 0)
    )
```

```{r pair4}
#| warning: false
ggpairs(abortion_data, 
  columns = c("postma4", "cai", "trustmostpeople", "godimportant", 
               "ideology", "satisfinancial"),
  columnLabels = c("postmaterialism", "child autonomy", "trust", "god", 
               "ideology", "satisfinancial")) +
   theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(angle = 45, size = 10),
    strip.text.y = element_text(angle = 0, hjust = 0)
    )
```

From these correlation matrices, we can conclude that the highest
correlations (above 0.2) are those between `godimportant` and `cai`
0.370 `godimportant` and `respectauthority` 0.244 `godimportant` and
`ideology` 0.243 `godimportant` and `nationalpride` 0.216
`nationalpride` and `age` 0.206 `nationalpride` and `respectauthority`
0.206 `respectauthority` and `cai` 0.205

We want to choose predictor variables that are not strongly correlated with each other, are relatively balanced in our data set, and visually appear to have a significant relationship with the outcome. Now that we have calculated the correlation matrix and know that some variables may have multicollinearity, we will select variables that do not exhibit significant multicollinearity for our model. Moreover, we can run several multiple linear regression models with different combinations of predictors to see which model performs best and is least affected by multicollinearity.

>>>>

We create a visualization and summary statistics for the response variable.

@@

We now extend our exploratory data analysis (EDA) to some predictor
variables of interest; namely, the year, ideology, Child Autonomy Index,
Importance of God, Respect for Authority and National Pride predictors.
The EDA for each variable comprises a histogram and a boxplot of the
response variable grouped by predictor value. Additionally, we have a
jitter plot to explore the potential for an interaction effect between
year and importance of God.

@@

@@

@@

@@

```{r correlations}
#| warning: false
ggpairs(abortion_data, 
  columns = c("year", "ideology", "cai", "godimportant", "respectauthority", 
              "nationalpride"),
  columnLabels = c("year", "ideology", "child autonomy", "imptnce of God", "resp for autho", 
              "natl pride")) +
   theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(angle = 45, size = 10),
    strip.text.y = element_text(angle = 0, hjust = 0)
    )
```

From these correlation matrices, we can conclude that the highest
correlations (above 0.2) are those between `godimportant` and `cai`
0.370 `godimportant` and `respectauthority` 0.244 `godimportant` and
`ideology` 0.243 `godimportant` and `nationalpride` 0.216
`nationalpride` and `respectauthority` 0.206 `respectauthority` and
`cai` 0.205

\pagebreak

## METHOD

We will conduct a Multiple Linear Regression (MLR) on the attitudes on
the justifiability of abortion against several other predictor variables
found in this data set. This is because our outcome variable, the
attitude towards abortion, is measured on a numeric scale from 1 to 10,
and there are multiple predictor variables that we feel are potentially
correlated with it.

We will now discuss the data cleaning and predictor selection process.
For observations, we removed the observations that were part of
"generational waves" before 1995, because many survey questions were
only added from 1995 onward, and observations before 1995 have a high
number of null values.

We removed the variable "wvsccode" as it is the same for all the
observations. The WVSC Code depends on the country the surveys were
conducted in, and all surveys in this data set were conducted in the
United States of America.

We chose to consider predictor variables that, from our exploratory data
analysis (EDA), looked like they were potentially correlated with the
outcome. Because the predictor variables were discrete, our EDA was
presented as a series of box plots. We observed the differences in
median and quartile values of abortion attitude across different
predictor values and ruled out variables with no visible hint of
relationship.

This led to our first set of predictor variables, which we designate as
Recipe1: 
- Year 
- Ideology 
- Child Autonomy Index 
- Importance of God 
- Respect for Authority 
- National Pride

We also wanted to choose predictor variables that are not strongly
correlated with each other, to avoid multicollinearity. From our EDA, we
found that several predictors had correlations of 0.2 or higher with
each other. Given these correlations, we decided to make a second recipe
(`Recipe2`) with the following predictor variables, after excluding
those with a high correlation with the `godimportant` variable:

-   Year
-   Importance of God

Finally, we considered that the distribution for the `godimportant`
variable is heavily left-skewed in our data set, which we observed in
the EDA. This may reduce the ability of a model that relies on the
"importance of God" predictor to explain the variations we observe in
the outcome. The same can be said of the `respectauthority` variable,
with more than 8 times the observations indicating "1" compared to the
observations indicating "-1".

Consequently, our third recipe (Recipe3) excludes `godimportant` and
uses the other variables that are not correlated with each other more
than 0.200:

-   Year
-   Ideology
-   Child Autonomy Index
-   National Pride

## Data Split

## Fit Model

Note: make sure to create recipes, specify engine, make workflows, and
fit models here

## Inference

\pagebreak

##RESULTS

## Model Evaluation

## Model Comparison

## Model Diagnosis

## Discussion of Influential Points
